```{r, include=FALSE}
base = read.csv("~/Downloads/base_data.csv")

base1 = base[which(base$year == c("1996")),]
base2 = base[which(base$year == c("2000")),]
base3 = base[which(base$year == c("2004")),]
base4 = base[which(base$year == c("2008")),]

base.total = rbind(base1,base2,base3,base4)
base.totalc = base.total[complete.cases(base.total),]
```
Summary of Medal Count:
```{r, echo=FALSE}
summary(base.totalc$count)
```
Distribution of Medal Count:
```{r, echo=FALSE}
hist(base$count)
```

Distribution of Medal Count in 2008:
```{r, echo=FALSE}
hist(base[which(base$year == c("2008")),c("count")])
```




GDP Summary:
```{r, echo=FALSE}
summary(base.totalc$gdp)
```
Population Summary:
```{r, echo=FALSE}
summary(base.totalc$pop)
```
Total Count of Communist / comm_soviet Countries:
```{r, echo=FALSE}
sum(base.totalc$comm_soviet)
```
Count of host, hosted within 8 years prior, or will be hosting within 8 years:
```{r, echo=FALSE}
sum(base.totalc$host)
```
Histograms of GDP
```{r, echo=FALSE}
par(mfrow=c(1,2))
#Histogram GDP
hist(base.totalc$gdp,main = "Untransformed",xlab = "GDP")
#Histogram log(GDP)
hist(log(base.totalc$gdp),main = "Transformed",xlab = "log(GDP)")
```
\
Histograms of Population:
```{r, echo=FALSE}
par(mfrow=c(1,2))
#Histogram Population
hist(base.totalc$pop,main = "Untransformed",xlab = "Population")
#Histogram log(Population)
hist(log(base.totalc$pop),main = "Transformed",xlab = "log(Population)")
```
\
Boxplot of $\text{log}(\text{GDP})$:
```{r, echo=FALSE}
#Log(GDP)
boxplot(log(base.totalc$gdp),ylab = "log(GDP)")
```
\
Boxplot of $\text{log}(\text{Population})$:
```{r, echo=FALSE}
#log(Population)
boxplot(log(base.totalc$pop),ylab = "log(Population)")
```

New Datafame to with logs:
```{r}
Olympic = base.totalc[,c(2,3,4,5,6,7,8)]
Olympic[,4:5] = log(Olympic[4:5],2)
```




\
Scatterplot Matrix of Variables:
```{r, echo=FALSE}
pairs(Olympic[,c(2,4,5,6,7)])
```
\
Correlation Matrix of Predictor Variables and Response Variable:
```{r, echo=FALSE}
cor(Olympic[,c(2,4,5,6,7)])
```




Poisson Model:
```{r}
Olympic.pois = glm(count~gdp+pop+comm_soviet+host,data = Olympic,family = poisson)
summary(Olympic.pois)
```

